---
title       : Skewed Normal Distributions
subtitle    : explaining some basics
author      : Hartwig Toedter
job         : Programmer
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
logo        : boss.jpg
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax]     # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Theoretical background
Given a random variable $X$, which has the probability denisity function
$$
\begin{equation}
f(x) = 2 \phi(x)\ \Phi(\alpha x)
\end{equation}
$$
where $\alpha$ is a number and
$$
\begin{equation}
\phi(x)= \exp(-x^2/2)/\sqrt{2\pi} ,
   \qquad \Phi(\alpha x) = \int_{-\infty}^{\alpha  x}\phi(t)\,dt
\end{equation}
$$
Set $\begin{equation}Y \: =\: \xi\, +\, \omega\, X\end{equation}$ and $Y$ is said to 
have a skewed normal distribution with parameters $\xi$, $\omega$ and $\alpha$
$$
\begin{equation}
Y\: \sim\:  SN(\xi,\,\omega^2,\,\alpha).
\end{equation}
$$
$\xi$ is called location, $\omega$ scale and $\alpha$ shape. In case $\alpha=0$ we get $Y\: \sim\: N(\xi,\,\omega^2)$ a normal distribution with mean $\xi$ and standard deviation $\omega$.

--- 

## Example of a left skewed distribution

```{r, echo=FALSE}
x = seq(-20, 20, by=0.01)
location = 0
scale = 1
skewness = -3

library(sn, quietly = TRUE, warn.conflicts = FALSE)

set.seed(1234)
data = rsn(1000, xi = location, omega = scale, alpha = skewness)
hist(data, breaks = 20, probability = TRUE)
lines(density(data))

y = dsn(x, xi = location, omega = scale, alpha = skewness)
lines(x,y, lty=2, col="red")
```

--- 

## Example of a right skewed distribution

```{r, echo=FALSE}
location = 0
scale = 1
skewness = 5

set.seed(4321)

data = rsn(1000, xi = location, omega = scale, alpha = skewness)
hist(data, breaks = 20, probability = TRUE)
lines(density(data))

y = dsn(x, xi = location, omega = scale, alpha = skewness)
lines(x,y, lty=2, col="red")
```

---

## References

For more information skew normals distributions and the sn R package see [The Skew-Normal Probability Distribution](http://azzalini.stat.unipd.it/SN/).
The source code

